---
title: "autumn-group-project-arya-jasmina-neenam"
author: "Rai N, Rana J, Shinde A"
date: "2023-12-21"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, results = 'FALSE'}
#LOAD PACKAGES
library(tidyverse)
library(janitor)
library(ggplot2)
library(lubridate)
library(dplyr)
library(grid)
library(patchwork)

#___________________________----
# IMPORT DATA ----
covid <- read_csv("covid_example_data (1)(1).csv")
head(covid)

#__________________________----
#CLEAN DATA----
covid <- janitor::clean_names(covid) # cleans the column names

colnames(covid) # checks the new variable names

covid <- rename(covid,
                "date_of_entry_for_case"="reprt_creationdt_false",
                "date_of_birth"="case_dob_false",
                "symptomatic_status"="contact_id" ,
                "date_of_first_symptoms"="sym_startdt_false",
                "symptoms_now_resolved"="sym_resolveddt_false",
                "date_of_hospital_admission"="hosp_admidt_false",
                "date_of_hospital_discharge"="hosp_dischdt_false",
                "date_of_death"= "died_dt_false",
                "date of positive PCR test" = "pos_sampledt_false",
                "case_ethnicity" = "case_eth") #re-names columns to easily understandable text

glimpse(covid) # displays column names in console

colnames(covid) # check for duplicate rows in the data


duplicated(covid)# produces a list of TRUE/FALSE statements for duplicated or not
sum() # sums all the TRUE statements   

covid <- covid[!(is.na(covid$case_age)),] # omits NA from case_age data

covid <- covid[!(is.na(covid$died_covid)),] #omits NA from died_covid data

covid <- covid[!(is.na(covid$case_gender)),]#omits NA from case_gender data

#___________________________----
#COMPARATIVE GRAPH OF AGE AND GENDER OF PATIENTS STUDIED----


pal <- c("tomato2","palegreen3", "lightgrey")

p1 <- covid %>% 
  ggplot(aes(x = case_gender, # gender displayed on X-axis
             y = case_age, # age displayed on Y-Axis
             fill = case_gender,
             colour = case_gender))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+ 
  scale_colour_manual(values = pal)+
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs(
    x = "pateint gender", #x-axis title
    y = "patient age", #y-axis title
    title = "Age verses gender of patients studied", # graph title
    subtitle = "Box and violin plot of age by gender")# graph sub-title


#___________________________----
#COMPARATIVE GRAPH OF AGE AND DATE TO SHOW DEATH RATE----

pal <- c("orange", "lightgrey", "cornflowerblue")

p2 <- covid %>% 
  ggplot(aes(x = died_covid,# weather they died or not displayed on x-axis
             y = case_age, # age of patient recorded
             fill = died_covid,
             colour = died_covid))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs(
    x = "death of patient",
    y = "patient age",
    title = "Age of patient death from covid",
    subtitle = "Box and violin plot of age by death rate") 

#__________________________----
#COMBINING PLOTS INTO ONE FIGURE----

(p1+p2)+
plot_layout(guides = "collect") #combines plot 1 and 2 in one figure
```

Figure 1. Plot A shows the roughly even distribution of ages of male and female patients in the data set. Plot B is a comparative plot to show the effect that a patients age has on their death rate, with a distribution being showing that the older the patient, the higher the mortality rate.These two figures provide an overview of the data set as a whole. 


```{r, results = 'FALSE'}
#LOAD PACKAGES
library(tidyverse)
library(ggplot2)
library(dplyr)
library(patchwork)

#___________________________----
# IMPORT DATA ----
covid <- read_csv("covid_example_data (1)(1).csv")
head(covid)


#__________________________----
#CREATING OBJECT----

symptoms_and_ages <- select(.data = covid, 
                            case_age, case_race, sym_fever, sym_myalgia, sym_losstastesmell, sym_sorethroat, sym_cough, sym_headache) # Creating a data set with all the columns I am interested in

symptpms_and_ages_omit  <- na.omit(symptoms_and_ages) # getting rid of missing variables in the data set 

symptpms_and_ages_omit_no_unknowns  <- filter(symptpms_and_ages_omit, 
                                              sym_fever != "Unk", 
                                              sym_myalgia != "Unk",
                                              sym_losstastesmell != "Unk",
                                              sym_sorethroat != "Unk",
                                              sym_cough != "Unk",
                                              sym_headache != "Unk") # getting rid of the rows containing unknown variables in the data set 
  

#__________________________----
#COMPARATIVE GRAPHS OF AGE AND PREVELANCE OF CERTAIN SYMPTOMS----

pal <- c("tomato2","cornflowerblue") #colour for graphs

#Graph for age and the presence of fever

p1 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_fever,
             y = case_age,
             fill = sym_fever,
             colour = sym_fever))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of fever",
        y = "Patients age",
        title = "Age of patients with symptoms of fever",
        subtitle = "Box and violin plot of the presence of fever by patients age")



#Graph for age and the presence of myalgia
p2 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_myalgia,
             y = case_age,
             fill = sym_myalgia,
             colour = sym_myalgia))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+# adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of myalgia",
        y = "Patients age",
        title = "Age of patients with symptoms of myalgia",
        subtitle = "Box and violin plot of the presence of myalgia by patients age")



#Graph for age and the presence of the loss of taste and smell
p3 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_losstastesmell,
             y = case_age,
             fill = sym_losstastesmell,
             colour = sym_losstastesmell))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of the loss of taste and smell",
        y = "Patients age",
        title = "Age of patients with symptoms of loss of taste and smell",
        subtitle = "Box and violin plot of the presence of the loss of taste and smell by patients age")


#Graph for age and the presence of a sore throat
p4 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_sorethroat,
             y = case_age,
             fill = sym_sorethroat,
             colour = sym_sorethroat))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of a sore throat",
        y = "Patients age",
        title = "Age of patients with symptoms of sore throat",
        subtitle = "Box and violin plot of the presence of a sore throat by patients age")



#Graph for age and the presence of cough
p5 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_cough,
             y = case_age,
             fill = sym_cough,
             colour = sym_cough))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of coughing",
        y = "Patients age",
        title = "Age of patients with symptoms of coughing",
        subtitle = "Box and violin plot of the presence of coughing by patients age")



#Graph for age and the presence of headache
p6 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_headache,
             y = case_age,
             fill = sym_headache,
             colour = sym_headache))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of headache",
        y = "Patients age",
        title = "Age of patients with headache symptoms",
        subtitle = "Box and violin plot of the presence of headache by patients age")


# merging all the plots together
(p1+p2+p3+p4+p5+p6)+
  plot_layout(guides = "collect")
```

Figure 2. 6 Plots showing the prevalence of different symptoms at different ages using patients who have been tested for all symptoms.

```{r, results = 'FALSE'}
#LOAD PACKAGES
library(tidyverse)
library(janitor)
library(ggplot2)
library(dplyr)
library(ggplot2)


#__________________________----
# ----IMPORT DATA ----
covid <- read_csv("covid_example_data (1)(1).csv")
head(covid)
#__________________________----
# ----CLEAN DATA----
covid <- janitor::clean_names(covid) # clean the column names

colnames(covid) # check the new variable names


#__________________________----
# ----SELECT DATA SETS REQUIRED----
symptomatic_and_case_race <- select(.data = covid, 
                                    case_race, symptomatic_status) 
#__________________________----
# ----REMOVE MISSING VARIABLES----
symptomatic_and_case_race_omit  <- na.omit(symptomatic_and_case_race) 

#__________________________----
# ----REMOVE UNKNOWN VALUES----
symptomatic_and_case_race_omit_no_unknown   <- filter(symptomatic_and_case_race_omit , 
                                                      case_race != "UNKNOWN")

#__________________________----
# ----CALCULATE OBSERVATION COUNT ACROSS SYMPTOMATIC STATUS----
symptomatic_and_case_race_omit_no_unknown %>% count(symptomatic_status)
symptomatic_and_case_race_omit_no_unknown %>% count(symptomatic_status, sort = TRUE)
symptomatic_and_case_race_omit_no_unknown %>% count(symptomatic_status, case_race, sort = TRUE)

#__________________________----
# ----ASSIGN THE COLOURS FOR THE KEY, LABEL THE KEY AND RACES----
colors = c("#009E73", "#CC79A7", "#56B4E9")

#__________________________----
# ----LABEL THE KEY AND RACES----
race <- c("Black", "White", "Asian", "Native American/Alaska native", "Pacific islander", "Other")
symptomatic_status <- c("Asymptomatic", "Symptomatic", "Unknown")

#__________________________----
# ----CREATE MATRIX----
#the matrix creates a data structure
Values <- matrix(c( 3721, 2693, 281, 5, 3, 474, 18143, 17215, 1737, 50, 41, 2802, 545, 355, 33, 5, 3, 53),
                 nrow = 3, ncol = 6, byrow = TRUE) #nrow tells us there are 3 rows (symptomatic_status) and ncol tells us there are 6 columns (race)

#__________________________----
# ----CREATE BAR CHART----
barplot(Values, main = "The distribution of symptomatic status across race", names.arg = ethnicity,
        xlab = "Race", ylab = "Number of Observations",
        #set limits and arrange font sizes
        ylim=c(0,20000),
        cex.name=0.50,
        col = colors, beside = TRUE)

#__________________________----
# ----ADD LEGEND FOR THE KEY----
legend("topright", symptomatic_status, cex = 1, fill = colors)

```
Figure 3. The figure demonstrates the distribution of the symptomatic status of COVID-19 across the six populations of race (Black, White,Asian,Native American/Alaska native, Pacific Islander and Other). The symptomatic status was measured in number of observations across a population.

**Reference List**



(1) Chang, W. (2024) R graphics cookbook, 2nd edition, 3.3 Making a Bar Graph of Counts. Available at: https://r-graphics.org/recipe-bar-graph-counts (Accessed: 08 January 2024). 

(2) Holtz, Y. (no date) Basic barplot with GGPLOT2, – the R Graph Gallery. Available at: https://r-graph-gallery.com/218-basic-barplots-with-ggplot2.html (Accessed: 08 January 2024). 

figure 1. (plot 1 + 2)
Leftwich, P.T. (no date) 5023Y:data science for biologists, Chapter 3 loading data,Chapter 4 data wrangling part 1, Chapter 5 data wrangling part 2, Chapter 6 Data visualization with ggplot 2, Chapter 7 Markdown. Available at: https://ueabio.github.io/data-sci-v1/markdown.html (Accessed: 08 January 2024).

(3) Leftwich, P.T. (no date) Chapter 7 Markdown: 5023Y, Chapter 7 Markdown | 5023Y. Available at: https://ueabio.github.io/data-sci-v1/markdown.html (Accessed: 08 January 2024). 

(4) R.K.- (2017) Bar plots, Quick-R: Bar Plots. Available at: https://www.statmethods.net/graphs/bar.html (Accessed: 08 January 2024). 