---
title: "autumn-group-project-arya-jasmina-neenam"
author: "Rai N, Rana J, Shinde A"
date: "2023-12-21"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, results = 'FALSE'}
#LOAD PACKAGES
library(tidyverse)
library(janitor)
library(ggplot2)
library(lubridate)
library(dplyr)
library(maps)
library(grid)
library(patchwork)

covid <- read_csv("covid_example_data (1)(1).csv")
head(covid)

covid <- covid[!(is.na(covid$case_age)),] # omits NA from case_age data

covid <- covid[!(is.na(covid$died_covid)),] #omits NA from died_covid data

covid <- covid[!(is.na(covid$case_gender)),]#omits NA from case_gender data

#___________________________----
#COMPARATIVE GRAPH OF AGE AND GENDER OF PATIENTS STUDIED----


pal <- c("tomato2","palegreen3", "lightgrey")

p1 <- covid %>% 
  ggplot(aes(x = case_gender, # gender displayed on X-axis
             y = case_age, # age displayed on Y-Axis
             fill = case_gender,
             colour = case_gender))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+ 
  scale_colour_manual(values = pal)+
  coord_flip()+ # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs(
    x = "pateint gender", #x-axis title
    y = "patient age", #y-axis title
    title = "Age verses gender of patients studied", # graph title
    subtitle = "Box and violin plot of age by gender")# graph sub-title


#___________________________----
#COMPARATIVE GRAPH OF AGE AND DATE TO SHOW DEATH RATE----

pal <- c("orange", "lightgrey", "cornflowerblue")

p2 <- covid %>% 
  ggplot(aes(x = died_covid,# weather they died or not displayed on x-axis
             y = case_age, # age of patient recorded
             fill = died_covid,
             colour = died_covid))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+ # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs(
    x = "death of patient",
    y = "patient age",
    title = "Age of patient death from covid",
    subtitle = "Box and violin plot of age by death rate") 

#__________________________----
#COMBINING PLOTS INTO ONE FIGURE----

(p1+p2)+
plot_layout(guides = "collect") #combines plot 1 and 2 in one figure
```

Figure 1. Plot A shows the roughly even distribution of ages of male and female patients in the data set. Plot B is a comparative plot to show the effect that a patients age has on their death rate, with a distribution being showing that the older the patient, the higher the mortality rate.

```{r, results = 'FALSE'}
#LOAD PACKAGES
library(tidyverse)
library(ggplot2)
library(dplyr)
library(patchwork)

#___________________________----
# IMPORT DATA ----
covid <- read_csv("covid_example_data (1)(1).csv")
head(covid)


#__________________________----
#CREATING OBJECT----

symptoms_and_ages <- select(.data = covid, 
                            case_age, case_race, sym_fever, sym_myalgia, sym_losstastesmell, sym_sorethroat, sym_cough, sym_headache) # Creating a data set with all the columns I am interested in

symptpms_and_ages_omit  <- na.omit(symptoms_and_ages) # getting rid of missing variables in the data set 

symptpms_and_ages_omit_no_unknowns  <- filter(symptpms_and_ages_omit, 
                                              sym_fever != "Unk", 
                                              sym_myalgia != "Unk",
                                              sym_losstastesmell != "Unk",
                                              sym_sorethroat != "Unk",
                                              sym_cough != "Unk",
                                              sym_headache != "Unk") # getting rid of the rows containing unknown variables in the data set 
  

#__________________________----
#COMPARATIVE GRAPHS OF AGE AND PREVELANCE OF CERTAIN SYMPTOMS----

pal <- c("tomato2","cornflowerblue") #colour for graphs

#Graph for age and the presence of fever

p1 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_fever,
             y = case_age,
             fill = sym_fever,
             colour = sym_fever))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of fever",
        y = "Patients age",
        title = "Age of patients with symptoms of fever",
        subtitle = "Box and violin plot of the presence of fever by patients age")



#Graph for age and the presence of myalgia
p2 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_myalgia,
             y = case_age,
             fill = sym_myalgia,
             colour = sym_myalgia))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+# adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of myalgia",
        y = "Patients age",
        title = "Age of patients with symptoms of myalgia",
        subtitle = "Box and violin plot of the presence of myalgia by patients age")



#Graph for age and the presence of the loss of taste and smell
p3 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_losstastesmell,
             y = case_age,
             fill = sym_losstastesmell,
             colour = sym_losstastesmell))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of the loss of taste and smell",
        y = "Patients age",
        title = "Age of patients with symptoms of loss of taste and smell",
        subtitle = "Box and violin plot of the presence of the loss of taste and smell by patients age")


#Graph for age and the presence of a sore throat
p4 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_sorethroat,
             y = case_age,
             fill = sym_sorethroat,
             colour = sym_sorethroat))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of a sore throat",
        y = "Patients age",
        title = "Age of patients with symptoms of sore throat",
        subtitle = "Box and violin plot of the presence of a sore throat by patients age")



#Graph for age and the presence of cough
p5 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_cough,
             y = case_age,
             fill = sym_cough,
             colour = sym_cough))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of coughing",
        y = "Patients age",
        title = "Age of patients with symptoms of coughing",
        subtitle = "Box and violin plot of the presence of coughing by patients age")



#Graph for age and the presence of headache
p6 <- symptpms_and_ages_omit_no_unknowns %>% 
  ggplot(aes(x = sym_headache,
             y = case_age,
             fill = sym_headache,
             colour = sym_headache))+
  geom_violin(alpha = 0.2)+
  geom_boxplot(width = 0.2,
               alpha = 0.7)+ # adds scale
  scale_fill_manual(values = pal)+
  scale_colour_manual(values = pal)+
  coord_flip()+  # flips axis direction
  theme_light()+ # adds in 'light' theme for cohesion between plots
  theme(legend.position = "none")+
  labs( x = "The presence of headache",
        y = "Patients age",
        title = "Age of patients with headache symptoms",
        subtitle = "Box and violin plot of the presence of headache by patients age")


# merging all the plots together
(p1+p2+p3+p4+p5+p6)+
  plot_layout(guides = "collect")
```

Figure 2. 6 Plots showing the prevalence of different symptoms at different ages using patients who have been tested for all symptoms.


